---
title: "BEN's_version"
format: html
---

Ben's work goes here

# Task 1

1.  What is the purpose of the lapply() function? What is the equivalent purrr function?

- The `lapply()` function in base R is used to apply a function to every element of a list (or vector) and return the results as a list. It helps you avoid writing explicit for loops and makes your code more concise. The equivalent function in the `purrr` package is `map()`.

2.  Suppose we have a list called `my_list`. Each element of the list is a numeric data frame (all columns are numeric). We want use `lapply()` to run the code `cor(numeric_matrix, method = "kendall")` on each element of the list. Write code to do this below! (I’m really trying to ask you how you specify `method = "kendall"` when calling `lapply()`)

```{r, eval = FALSE}
custom_my_list <- lapply(X = my_list, FUN = cor, method = "kendall")
```

3.  What are two advantages of using purrr functions instead of the BaseR apply family?

- More consistent and readable syntax.

  - The `purrr` functions all follow a consistent naming convention (`map()`, `map_dbl()`, `map_chr()`, `map_df()`, etc.), making them easier to learn and use than the various Base R functions (`lapply()`, `sapply()`, `apply()`, etc.).

- Type-safe outputs.

  - Functions like `map_dbl()`, `map_int()`, `map_chr()`, and `map_lgl()` guarantee the type of the output. If the function cannot produce the requested type, `purrr` throws an error instead of silently simplifying the result, making code more predictable and easier to debug.

4.  What is a side-effect function?

- A side-effect function performs an action such as printing, writing files, or saving plots, rather than returning a value

5.  You can safely name a variable `sd` inside a function because of lexical scoping. When you assign a value to a variable, it is stored in the function's local environment, protecting it from built-in functions outside the function scope, such as the `sd()` function.

# Task 2

We want to write a `getRMSE()` function that will calculate the RMSE when given a *vector* of responses and a *vector* of predictions. We also want to allow for additional arguments in the `mean()` argument so that we can remove any `NA` values

```{r}
getRMSE <- function(response, prediction, ...) {  # ... will allow for additional arguments
  sqrt(mean((response - prediction)^2, ...))
}
```

Generate response and prediction vectors. We'll use these to test our function.

```{r}
# generate our prediction and response vectors
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

```{r}
# test the function
getRMSE(response = resp, prediction = pred)
```

Add some `NA` values to make sure our optional `...` argument works.

```{r}
resp[c(17, 49)] <- NA_real_  # 49 and 17 were chosen randomly. Feel free to change
```

Test the function.

```{r}
getRMSE(response = resp, prediction = pred, na.rm = TRUE) # make sure you specify na.rm = TRUE
```


Now lets write a function that will calculate MAE and call it `getMAE()`.

```{r}
getMAE <- function(response, prediction, ...) {
  mean(abs(response - prediction), ...)
}
```


Generate response and prediction vectors. We'll use these to test our function.

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```


Test the function.

```{r}
getMAE(response = resp, prediction = pred)
```

Replace two values in `resp` with `NA` to test our additional `...` argument

```{r}
resp[c(2, 37)] <- NA_real_  # again, 2 and 37 were chosen randomly by me
```


Test the function.

```{r}
getMAE(response = resp, prediction = pred, na.rm = TRUE)  # make sure to specify na.rm = TRUE
```


Now let's write a wrapper function that calls our helper functions, `getRMSE()` and `getMAE()`. By default, the function will return both metrics, but we can specify which metric(s) we want using a character vector. We'll also check that the inputs are numeric atomic vectors before proceeding.

```{r}
getMetrics <- function(response, prediction,
                       metrics = c("RMSE", "MAE"), ...) {

  # check that both inputs are numeric atomic vectors
  if (!(is.vector(response) && is.atomic(response) && is.numeric(response) &&
        is.vector(prediction) && is.atomic(prediction) && is.numeric(prediction))) {
    message("Both response and prediction must be numeric atomic vectors.")
    return(invisible(NULL))
  }

  results <- list()

  if ("RMSE" %in% metrics) {
    results$RMSE <- getRMSE(response, prediction, ...)
  }

  if ("MAE" %in% metrics) {
    results$MAE <- getMAE(response, prediction, ...)
  }

  return(results)
}
```

Generate response and prediction vectors.

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

Test the function by requesting each metric individually and then both.

```{r}
getMetrics(response = resp, prediction = pred, metrics = "RMSE")
getMetrics(response = resp, prediction = pred, metrics = "MAE")
getMetrics(response = resp, prediction = pred)
```

Now replace two response values with `NA`.

```{r}
resp[c(12, 84)] <- NA_real_
```

Test the function again.

```{r}
getMetrics(response = resp, prediction = pred, na.rm = TRUE)
```

Finally, test the function by passing incorrect data.

```{r}
bad_data <- data.frame(resp = resp)

getMetrics(response = bad_data, prediction = pred)
```

Great! Everything seems to work.