---
title: "BEN's_version"
format: html
---

Ben's work goes here

# Task 1

1.  What is the purpose of the lapply() function? What is the equivalent purrr function?

- The `lapply()` function in base R is used to apply a function to every element of a list (or vector) and return the results as a list. It helps you avoid writing explicit for loops and makes your code more concise. The equivalent function in the `purrr` package is `map()`.

2.  Suppose we have a list called `my_list`. Each element of the list is a numeric data frame (all columns are numeric). We want use `lapply()` to run the code `cor(numeric_matrix, method = "kendall")` on each element of the list. Write code to do this below! (I’m really trying to ask you how you specify `method = "kendall"` when calling `lapply()`)

```{r, eval = FALSE}
custom_my_list <- lapply(X = my_list, FUN = cor, method = "kendall")
```

3.  What are two advantages of using purrr functions instead of the BaseR apply family?

- More consistent and readable syntax.

  - The `purrr` functions all follow a consistent naming convention (`map()`, `map_dbl()`, `map_chr()`, `map_df()`, etc.), making them easier to learn and use than the various Base R functions (`lapply()`, `sapply()`, `apply()`, etc.).

- Type-safe outputs.

  - Functions like `map_dbl()`, `map_int()`, `map_chr()`, and `map_lgl()` guarantee the type of the output. If the function cannot produce the requested type, `purrr` throws an error instead of silently simplifying the result, making code more predictable and easier to debug.

4.  What is a side-effect function?

- A side-effect function performs an action such as printing, writing files, or saving plots, rather than returning a value

5.  You can safely name a variable `sd` inside a function because of lexical scoping. When you assign a value to a variable, it is stored in the function's local environment, protecting it from built-in functions outside the function scope, such as the `sd()` function.

# Task 2

We want to write a `getRMSE()` function that will calculate the RMSE when given a *vector* of responses and a *vector* of predictions. We also want to allow for additional arguments in the `mean()` argument so that we can remove any `NA` values

```{r}
getRMSE <- function(response, prediction, ...) {  # ... will allow for additional arguments
  sqrt(mean((response - prediction)^2, ...))
}
```

Now lets run this code to test our function

```{r}
# generate our prediction and response vectors
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

```{r}
# test the function
getRMSE(response = resp, prediction = pred)
```

Add some `NA` values to make sure our optional `...` argument works

```{r}
resp[c(17, 49)] <- NA_real_  # 49 and 17 were chosen randomly. Feel free to change
```

Test the function

```{r}
getRMSE(response = resp, prediction = pred, na.rm = TRUE) # make sure you specify na.rm = TRUE
```

