---
title: "ST 558 HW 5 - Final Version"
error: true
format: 
  html:
    toc: true
    theme: cosmo
---

# Task 1: Conceptual Questions:

1. `lapply()` applies a function to every element in a list/vector/data frame and returns a list. It helps you avoid writing explicit for loops and makes your code more concise. Equivalently, we could use `map()` from the `purrr` package.
2. If we call `?lapply`, we can see that it's called with `lapply(X, FUN, ...)`, where `...` is optional arguments passed to the function `FUN`. Hence, we'd pass the following:
```{r, eval = FALSE}
lapply(numeric_matrix, cor, method = "kendall")
```

3. Compared to baseR `xapply()` functions, `purrr` functions:

- Offer greater **consistency in syntax** regarding the selection and order of arguments each function takes.
- Have **less ambiguity in the type of object returned**; baseR functions may return different types based on the type of the input vector (like for `sapply()`).
4. A side effect function is a function which **doesn't simply return a transformation of the inputs**. They may have no return at all. For example, the plot function takes data as an input and returns a plot object.
5. **Lexical scoping** prevents variable names within functions from conflicting with built-in functions. When functions are executed in R, a temporary environment is created. This allows the use of object names that would ordinarily cause problems, because the conflicting functions are (generally) not part of the temporary environment.

# Task 2: Writing Functions
## RMSE Function:
We just need to validate the user's input and return the RMSE. The elipses `...` allow for the user to pass additional arguments to the `mean()` function.
```{r}
getRMSE <- function(response, predicted, ...) {
  # data validation
  if(!is.vector(response)) stop("Error: Reponse isn't a vector")
  if(!is.atomic(response)) stop("Error: Reponse isn't an atomic vector")
  if(!is.numeric(response)) stop("Error: Response isn't a numeric vector")
  if(!is.vector(predicted)) stop("Error: Predicted isn't a vector")
  if(!is.atomic(predicted)) stop("Error: Predicted isn't an atomic vector")
  if(!is.numeric(predicted)) stop("Error: Predicted isn't a numeric vector")
  if(length(response) != length(predicted)) stop("Error: Reponse and prediction vectors have unequal lengths")
  
  # return rmse
  sqrt(mean((response - predicted)**2, ...))
}
```

## Generating RMSE Test Data:
Let's run the given code to generate test data:
```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

And test the RMSE function with it:
```{r}
getRMSE(resp, pred)
```
Seems legit. We can replace a few values with `NA` and test passing the `na.rm = TRUE` argument:
```{r}
missing_idx <- sample(length(resp), size = 2)
resp[missing_idx] <- NA_real_

getRMSE(resp, pred)
getRMSE(resp, pred, na.rm = TRUE)
```
Great! It looks like we can pass arguments to the `mean()` function!

## MAE Function
The `getMAE()` function is very similar to the previous function:
```{r}
getMAE <- function(response, predicted, ...) {
  # data validation
  if(!is.vector(response)) stop("Error: Reponse isn't a vector")
  if(!is.atomic(response)) stop("Error: Reponse isn't an atomic vector")
  if(!is.numeric(response)) stop("Error: Response isn't a numeric vector")
  if(!is.vector(predicted)) stop("Error: Predicted isn't a vector")
  if(!is.atomic(predicted)) stop("Error: Predicted isn't an atomic vector")
  if(!is.numeric(predicted)) stop("Error: Predicted isn't a numeric vector")
  if(length(response) != length(predicted)) stop("Error: Reponse and prediction vectors have unequal lengths")
  
  # return mae
  sqrt(mean(abs(response - predicted), ...))
}
```

Testing it:
```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

getMAE(resp, pred)
```
Sweet, now with missing data:
```{r}
missing_idx <- sample(length(resp), size = 2)
resp[missing_idx] <- NA_real_

getRMSE(resp, pred)
getRMSE(resp, pred, na.rm = TRUE)
```
Looking good!

## Wrapper Function
Last, we write a wrapper function to call one or both of the helper functions we just wrote:
```{r}
getMetric <- function(resp, pred, metrics, ...) {
  valid_metrics <- c("RMSE", "MAE")
  
  if(length(setdiff(metrics, valid_metrics)) > 0) {
    stop("Error: Invalid metrics specified. Specify at least one of: RMSE, MAE")
  }
  
  output <- numeric()
  
  if("RMSE" %in% metrics) {
    output["RMSE"] <- getRMSE(resp, pred, ...)
  }
  if("MAE" %in% metrics) {
    output["MAE"] <- getMAE(resp, pred, ...)
  }
  
  output
}
```

Again, generate data to test:
```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

Test querying metrics separately and together:
```{r}
getMetric(resp, pred, "MAE")
getMetric(resp, pred, "RMSE")
getMetric(resp, pred, c("MAE", "RMSE"))
```
Let's add some missingness and retest:
```{r}
missing_idx <- sample(length(resp), size = 2)
resp[missing_idx] <- NA_real_

getMetric(resp, pred, c("MAE", "RMSE"))
getMetric(resp, pred, c("MAE", "RMSE"), na.rm = TRUE)
```
Excellent! Lastly, we should see what happens if we pass a data frame instead of a vector:
```{r}
resp_df <- data.frame(resp)

getMetric(resp_df, pred, "MAE")
```
Sweet! The data validation we did earlier still works when called in the wrapper function.
